IPv6 over Low-Power Wireless Personal Area Networks\\ MSc. Thesis Proposal

### 2009-12-18 trunk/doc/msc-proposal/msc-nd.tex
proposa: ND

\section{Bootstrapping and Basic Operation}
Bootstrapping of a LoWPAN node consists of several steps.  At first, a node is required to autoconfigure at least one address, a link-local address, which is derived from the IEEE 64-bit extended MAC. In order to receive RAs from routers, a node joins the all-nodes multicast address and, if the node is a router, the all-routers multicast address. Once the interfaces have been initialized, a node listens for Router Advertisements (RA) from Edge Routers or LoWPAN Routers, or broadcasts a Router Solicitation (RS). Upon receipt of the RA, the node forms an optimistic global unique address with stateless address autoconfiguration and chooses one or more default routers (routers are present only in a Route Over configuration).

The constructed global address is tentative as long as the binding is not confirmed by the ER. To accomplish this, the node performs initial registration with the ER by sending a unicast Node Registration (NR) message with a list of all addresses it wants to register. The ER replies with a Node Confirmation(NC), which includes the set of addresses confirmed to be bound to the Whiteboard of the ER. Once the node has received the NC message, it is capable to send packets to any IPv6 address inside or outside the LoWPAN. Each router maintains a set of information about nodes that are currently registered through it, called the binding table. Nodes send periodic registration messages in order to maintain their bindings in the router binding tables and Edge Router Whiteboard. The detection of duplicate addresses (DAD) is also performed as part of the node registration process by the ER using a lookup on the Whiteboard.

This information about link-local addresses is collected during the node registration process. Nodes store the information about router link-local addresses in their default router list, while routers keep the information about nodes in their binding tables.

\section{Message Formats}

6LoWPAN-ND makes use of Router Solicitation (RS)/Router Advertisement (RA) message exchanges similar to classic ND. In 6LoWPAN-ND RA messages may carry additional options for context dissemination and are reduced in size. In addition to RS and RA messages, 6LoWPAN-ND defines two new ICMP packet types: Node Registration (NR), which is sent by a node to an Edge Router to register a binding for an IPv6 address in the Whiteboard, and Node Confirmation (NC) by which an Edge Router replies to the registering node.

>>>>

\section{Bootstrapping and Basic Operation}
Bootstrapping of a LoWPAN node consists of several steps.  At first, a node is required to autoconfigure at least one address, a link-local address, which is derived from the IEEE 64-bit extended MAC. In order to receive RAs from routers, a node joins the all-nodes multicast address and, if the node is a router, the all-routers multicast address. Once the interfaces have been initialized, a node listens for Router Advertisements (RA), or broadcasts a Router Solicitation (RS). Upon receipt of the RA, the node forms an optimistic global unique address with stateless address autoconfiguration and chooses one or more default routers (routers are present only in a Route Over configuration).

The constructed global address is tentative as long as the binding is not confirmed by the ER. To accomplish this, the node performs initial registration with a link-local ER  or LoWPAN Router by sending a unicast Node Registration (NR) message with a list of all addresses it wants to register. Registering directly with an ER is preferred, although all LoWPAN Routers have the ability to relay NR/NC messages on behalf of a node to the ER. The ER replies with a Node Confirmation(NC), which includes the set of addresses confirmed to be bound to the Whiteboard of the ER. Once the node has received the NC message, it is capable to send packets to any IPv6 address inside or outside the LoWPAN. Each router maintains a set of information about nodes that are currently registered through it, called the binding table. Nodes send periodic registration messages in order to maintain their bindings in the router binding tables and Edge Router Whiteboard. The detection of duplicate addresses (DAD) is also performed as part of the node registration process by the ER using a lookup on the Whiteboard.

Next-hop determination assumes destinations are non-local and forwarded to a default router except for link-local scope addresses. The information about link-local addresses is collected during the node registration process. Nodes store router link-local addresses in their default router list, whereas routers resolve the link-layer addresses of the attached nodes by looking up on the binding table.

\section{Message Types}

6LoWPAN-ND makes use of Router Solicitation (RS) and Router Advertisement (RA) message exchanges in a similar way to classic ND. However, 6LoWPAN-ND RA messages may carry additional options for context dissemination and are reduced in size. In addition to RS and RA messages, 6LoWPAN-ND defines two new ICMP packet types, Node Registration (NR) and Node Confirmation (NC), used by a node to register with an ER, and by the ER to confirm the
binding.

\subsection{Router Solicitation Message}
The format of RS message for 6LoWPAN is identical to the classic RS message \cite{rfc4861}. If a node has only the addresses that are not yet confirmed by an ER, then the IPv6 unspecified address is used as the IPv6 source address. The Source Link-Layer Address Option \cite{rfc4861} used in IPv6 ND is not included in the RSs and the Owner Interface Identifier Option, described in Section \ref{sec:nd.oiio}, is used instead.

\subsection{Router Advertisement Message}
The format of RA message for 6LoWPAN is identical to the classic \cite{rfc4861} RA message. However, the use of a 2-bit  Default Router Preference (Prf) flag \cite{rfc4191} is defined in the context of 6LoWPAN. LoWPAN Edge Routers set Prf to $01$ indicating their high preference, whereas LoWPAN Routers with and without ER availability use the values $00$ and $01$ respectively for normal and low preference. RAs are sent either to all-nodes multicast or to a link-local unicast address as a response to an RS. RA messages use additional 6LoWPAN Information and 6LoWPAN Summary Options described in Sections \ref{nd.option.info} and \ref{nd.option.summary} respectively.

\subsection{Neighbor Solicitation/Advertisement Message}
Neighbor Solicitation(NS) and Neighbor Advertisement(NA) messages are employed only between ERs on the backbone link in Extended LoWPAN. Extended LoWPAN (shown in Figure \ref{fig:ext.lowpan})) allows to form a single subnet out of multiple LoWPANs interconnected by a backbone link via Edge Routers. The format of NS/NA messages is identical to the classic NS/NA messages \cite{rfc4861} with additional use of Owner Interface Identifier Option, described in Section \ref{sec:nd.oiio}.

\begin{figure}[htp]
\begin{mylisting}
\begin{verbatim}
               +-----+                 +-----+
               |     | Router          |     | Host
               |     |                 |     |
               +-----+                 +-----+
                  |                       |
                  |     Backbone link     |
            +--------------------+------------------+
            |                    |                  |
         +-----+             +-----+             +-----+
         |     | Edge        |     | Edge        |     | Edge
         |     | Router      |     | Router      |     | Router
         +-----+             +-----+             +-----+
            o         o       o   o  o      o        o o
        o o   o  o  o  o  o o   o  o  o  o  o   o  o  o  o
       o  o o  o o   o    o   o  o  o  o     o   o  o  o o
       o   o  o  o     o    o    o  o     o      o  o   o
         o   o o     o          o  o      o    o       o
\end{verbatim}
\end{mylisting}
\caption{Extended LoWPAN}\label{fig:ext.lowpan}
\end{figure}

\subsection{Node Registration/Confirmation Message}
In order to register for the first time, the node sends an NR message to the link-local unicast address of a local ER or LoWPAN Router. Since the node still has only an optimistic address, not yet confirmed by an Edge Router, the IPv6 unspecified address is used as the source address of the NR message. To renew bindings the subsequent NR messages use the link-local IPv6 address of the sender as the source address.

The message format for NR/NC messages is shown in Figure \ref{fig:nr.nc.format}. The values for the Type field will be defined in the future to distinguish Node Registration and Node Confirmation messages. The Status field is used only in NC messages and specifies the result of the registation procedure. A 4-bit unsigned Code provides additional details for the message, e.g. it may indicate whether the NR/NC is sent directly or relayed by a router. The Checksum field carries the ICMP checksum. A unique Transaction ID (TID) is used to match replies and a NC TID corresponds to the NR TID. Its value is incremented upon each new registration. A 1-bit Primary flag (P) indicates whether the ER is primary and may represent the node on the backbone in a Extended LoWPAN. A Router flag indicates the role of the node sending the NR message (0 is used by hosts and 1 by routers).  The Reserved field is unused and initialized to zero. The Binding Lifetime is the amount of time in minutes remaining before the binding of the owner interface identifier expires. The Advertising Lifetime field indicates the amount of time in units of 10 seconds the node will advertise itself to its local router using a NR local refresh. The Owner Nonce field is a 32-bit value generated randomly by the node upon booting.
And the Owner Interface Identifier (OII) is a globally unique identifier for the requesting host's interface.

A NR message includes a 6LoWPAN  Address Option, described in Section \ref{nd.option.address}, for each address the host wants to bind for the interface, and a NC message may carry a 6LoWPAN Information Option, described in Section \ref{nd.option.info}, containing information about the prefixes of the LoWPAN along with other context information.

\begin{figure}[htp]
\begin{mylisting}
\begin{verbatim}
    0                   1                   2                   3
    0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1
   +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
   |     Type      |Status | Code  |           Checksum            |
   +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
   |      TID      |P|R|                Reserved                   |
   +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
   |        Binding Lifetime       |     Advertising Interval      |
   +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
   |                         Owner Nonce                           |
   +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
   |                                                               |
   +                  Owner Interface Identifier                   +
   |                                                               |
   +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
   |   Registration option(s)...
   +-+-+-+-+-+-+-+-+-+-+-+-+-+
\end{verbatim}
\end{mylisting}
\caption{Node Registration/Confirmation message format}\label{fig:nr.nc.format}
\end{figure}

\subsection{Message Options}
\subsubsection{6LoWPAN Address Option}\label{nd.option.address}

\subsubsection{6LoWPAN Information Option}\label{nd.option.info}

\subsubsection{6LoWPAN Summary Option}\label{nd.option.summary}

\subsubsection{Owner Interface Identifier Option}\label{sec:nd.oiio}

### 2009-12-18 trunk/doc/msc-proposal/msc-proposal.tex
proposa: ND

%%
%	Transmission of IPv6 Packets
%%
\chapter{Transmission of IPv6 Packets over IEEE 802.15.4}\label{ch:6lowpan}
A low-power wireless personal area network (LoWPAN) is a simple low throughput wireless network comprising low cost and low power devices that conform to the IEEE 802.15.4-2003 standard. The IEEE 802.15.4-2003 standard specifies node-to-node frame delivery between the devices which are within reachable distance from each other. On top of IEEE 802.15.4 a network layer (layer 3 of the seven-layer OSI model) has to be defined that would provide end-to-end packet delivery including routing through intermediate hosts.

@@

\nocite{ieee802.15.4} 
\nocite{eui64} 
\nocite{rfc780}
\nocite{rfc2460}
\nocite{rfc2464}
\nocite{rfc2474}
\nocite{rfc3168}
\nocite{rfc3775}
\nocite{rfc4291} 
\nocite{rfc4443}
\nocite{rfc4861}
\nocite{rfc4862}
\nocite{rfc4919}
\nocite{rfc4944} 
\nocite{draft-usecases-05} 
\nocite{draft-hc-06} 
\nocite{draft-nd-07}

>>>>

%%
%	Transmission of IPv6 Packets
%%
\chapter{Transmission of IPv6 Packets over 802.15.4}\label{ch:6lowpan}
A low-power wireless personal area network (LoWPAN) is a simple low throughput wireless network comprising low cost and low power devices that conform to the IEEE 802.15.4-2003 standard. The IEEE 802.15.4-2003 standard specifies node-to-node frame delivery between the devices which are within reachable distance from each other. On top of IEEE 802.15.4 a network layer (layer 3 of the seven-layer OSI model) has to be defined that would provide end-to-end packet delivery including routing through intermediate hosts.

@@

\nocite{ieee802.15.4} 
\nocite{eui64} 
\nocite{rfc780}
\nocite{rfc2460}
\nocite{rfc2464}
\nocite{rfc2474}
\nocite{rfc3168}
\nocite{rfc3775}
\nocite{rfc4191}
\nocite{rfc4291} 
\nocite{rfc4443}
\nocite{rfc4861}
\nocite{rfc4862}
\nocite{rfc4919}
\nocite{rfc4944} 
\nocite{draft-usecases-05} 
\nocite{draft-hc-06} 
\nocite{draft-nd-07}
